---
output: pdf_document
---
#Chpater 3: R Codes Used in this Report   

##3.1 R codes used in Chapter 1  

###3.1.1 Graph of loss function for $\tau$ = 1/2 and $\tau$ = 1/3
```{r, eval = F}
par(mfrow = c(1, 2))
 plot.new()
 plot.window(xlim = c(-4, 4), ylim = c(-2, 2))
 axis(1, pos = 0, tick = F, labels = F)
 title(main = expression("Graph of loss function for " *  tau *" = 1/2"))
# axis(2, 0:2, pos = c(0, 0), tick = F)
abline(v = 0)
 axis(2, pos = 0, tick = F, labels = F)
 abline(h = 0)
 x1 <- seq(0, 4, length = 1000)
 y <- x1 * .5
 points(x1, y, type = "l", lwd = "2")
 x1 <- seq(-4, 0, length = 1000)
 y <- x1 * -.5
mtext(expression(rho[(1/2)](u)), c(-0.5, 3))
 points(x1, y , type = "l", lwd = "2")
 plot.new()
 plot.window(xlim = c(-4, 4), ylim = c(-2, 2))
 axis(1, pos = 0, tick = F, labels = F)
 # axis(2, 0:2, pos = c(0, 0), tick = F)
 abline(v = 0)
 axis(2, pos = 0, tick = F, labels = F)
 title(main = expression("Graph of loss function for " *  tau *" = 1/3"))
 abline(h = 0)
 x1 <- seq(0, 4, length = 1000)
 y <- x1 * (1/3)
 points(x1, y, type = "l", lwd = "2")
 x1 <- seq(-4, 0, length = 1000)
 y <- x1 * (-2/3)
 points(x1, y , type = "l", lwd = "2")
mtext(expression(rho[(1/3)](u)), c(-0.5, 3))
```
###3.1.2 To draw graphs of CDF and quantile function in discrete case.  

```{r,eval = F}
par(mfrow = c(1, 2))
 x <- c(1, 3/2, 2, 5/2, 3)
 Fx <- c(1/10, 1/5, 3/10, 3/10, 1/10)
x1 <- c(0, 1/10, 3/10, 6/10, 9/10)
x2 <- c(1/10, 3/10, 6/10, 9/10, 1)
# graph of CDF for the above data
 plot.new()
 plot.window(xlim = c(0, 3.1), ylim = c(0, 1.1))
 axis(1, at = c(1, 3/2, 2, 5/2, 3))
 axis(2, at = c(1/10, 3/10, 6/10, 9/10, 1))
 title(main = "Graph of CDF")
 points(x, x2, pch = 16)
 for(i in 1:(length(x2) - 1)){
     segments(x[i], x2[i], x[i+1], x2[i])
 }
segments(3, 1, 3.3, 1)
 box()
for(i in 1:length(x)){
    abline(v = x[i], lty = "dotted")
}
 
 # grpah of quantile function  for discrete distribution function
 
plot.new()
plot.window(xlim = c(0, 1.2), ylim = c(0, 3.5))
 axis(1, at = c(1/10, 3/10, 6/10, 9/10, 1))
 axis(2, at = c(1, 3/2, 2, 5/2, 3))
 points(x2, x, pch = 16)
 title(main = "Graph of Quantile Function")
 for(i in 1:(length(x1) - 1)){
     segments(x1[i], x[i], x1[i+1], x[i])
 }
segments(9/10, 3, 1, 3)
box()
for(i in 1:length(x2)){
    abline(v = x2[i], lty = "dotted")
}

 

```
###3.1.3 Graph of CDF and quantile function for continuous case.   
```{r,eval = F}
par(mfrow = c(1, 2))
 curve(x/(x+1), from = 0, to = 100, main = "Graph of CDF", xlab = "x", ylab = "CDF")
curve(x/(1 - x), from = 0, to = 1, main = "Graph of Quantile Funtion", xlab = "tau", ylab = "Q(tau)")

```   

##3.2 R codes used in chapter 2  
   
###3.2.1 Graph of CDF for two stochastic random variable  

```{r, eval =FALSE}
 # grpah for cdf of two stochastic random variable
 curve(1-exp(-x/2), from = 0, to = 10, col = "blue", ylab = "CDF")
curve(1-exp(-x/5), from = 0, to = 10, col = "red", add = T)
 segments(2, 0 , 2, (1 - exp(-1)))
segments(5, 0, 5, (1 - exp(-1)))
```
###3.2.2 Graph of scatterplot, OLS line and quantile regression line for various $\tau$.  
 
```{r, echo = TRUE, eval=FALSE}
library(quantreg)
data(engel)
# some rows of dataset
head(engel)
attach(engel)# to make variable accessible by it's name
# number of rows and columns
dimension <- dim(engel)
# to make a scatterplot and quantile regressor line
plot(income,foodexp,cex=.25,type="n",xlab="Household Income",
    ylab="Food Expenditure")
points(income,foodexp,cex=.5,col="blue")
abline(rq(foodexp~income,tau=.5),col="blue")
abline(lm(foodexp~income),lty=2,col="red") #the dreaded ols line
taus <- c(.05,.1,.25,.75,.90,.95)
f <- rq(foodexp ~ income, tau = taus)
for( i in 1:length(taus)){
    abline(coef(f)[,i],col="gray")
}
```
###3.2.3 Code for analysis of health data in sequence they appear in report.
```{r, echo=TRUE, eval=FALSE}

# reading data from csv file named as "quanitle_health.csv"
mydata<- read.csv("C:/Users/Ashvani/Documents/courseProject/quantile_health.csv")
attach(mydata) # to address data component without referring to data variable
head(mydata) # few data rows 
```

```{r, echo=TRUE,eval=FALSE}
Y <- cbind(totexp) # dependent variable
X <- cbind(suppins, totchr, age, female, white) # design matrix
# plotting percentiles of dependent variable
plot(x = seq(0, 1, by = .005),quantile(Y, seq(0, 1, by = .005)), type = "l", 
     col = "red", lwd = "2", xlab = "percentiles", ylab = "total medical expenditure")

```

```{r,fig.height=8, warning=FALSE, echo=TRUE, eval=FALSE}
# OLS regression
olsreg <- lm(Y ~ X, data=mydata)
summary(olsreg)
# Quantile regression
quantreg25 <- rq(Y ~ X, data=mydata, tau=0.25)
summary(quantreg25)
quantreg50 <- rq(Y ~ X, data=mydata, tau=0.5)
summary(quantreg50)
quantreg75 <- rq(Y ~ X, data=mydata, tau=0.75)
summary(quantreg75)
#quantreg90 <- rq(Y ~ X, data=mydata, tau=0.9)
# summary(quantreg90)
```


```{r, echo=TRUE, warning=FALSE, eval=FALSE}
# Plotting data
quantreg.all <- rq(Y ~ X, tau = seq(0.05, 0.95, by = 0.05), data=mydata)
quantreg.plot <- summary(quantreg.all)
plot(quantreg.plot)


```  