---
output: pdf_document
---
# Chapter 1 : Basics of Quantile Function  
## 1.1 Need for Quantile Regression
Linear regression model is most simple and popular model for statistical modeling. Suppose $X_1$, $X_2$, ............, $X_p$ are p-covariates and Y is the response variable then linear regression model is written as  
         $$
            \begin{aligned}
            E[Y|X = x] &= \beta_0 + \beta_1  X_1 + \beta_2  X_2 + ........... + \beta_p  X_p. 
            \end{aligned}
         $$
Hence linear regression estimates mean value of the response variable for given values of the covariates. There are situations where we want to see behaviour of response variable for left or right tail of covariates. Suppose we want to see how _infant birth weight_ is affected by covariates such as _the gender of the infant_, _prenatal care_ and _smoking status of mother during pregnancy_. Suppose linear regression model is  
$$ 
  \begin{aligned}
  Y &= \beta_0 + \beta_1*Boy + \beta_2*Prenatal Care + \beta_3*Smoke
  \end{aligned}
$$  
This model estimates how on average mothers characteristics affect the birth weights of infants. This model can answer significance of prenatal care i.e. "is prenatal care important?". But it can not answer "does prenatal care influence birth weight differently for infants with low birth weight than for high birth weight?". This question can be answered using *quantile regression*. To support importance of *quantile regression* **Mosteller and Tukey**  say,  
  
>What the regression curve does is give a grand summary for the averages of the
distributions corresponding to the set of x's.We could go further and compute
several different regression curves corresponding to the various percentage
points of the distributions and thus get a more complete picture of the set.
Ordinarily this is not done, and so regression often gives a rather incomplete
picture. Just as the mean gives an incomplete picture of a single distribution,
so the regression curve gives a correspondingly incomplete picture for a set
of distributions.  
  
In quantile regression conditional mean surface is replaced by conditional quantile surface.  
  
## 1.2 Quantiles
Let X be a real-valued random variable with distribution function F, then for 0 < $\tau$ < 1  
$$ 
   \begin{aligned}
    F^{-1}(\tau) &= \inf{\{x : F(x) \ge \tau\}}
    \end{aligned}
$$  
is called the **$\tau^{th}$ quantile** of X. The median, $F^{-1}$(1/2), plays the central role. 
  
### 1.2.1 Quantile as solution of an optimization problem
Consider a decision theoretic problem : a point estimate is required for a random variable with distribution function F. Suppose **loss function** is 
$$
   \begin{aligned}
   \rho_{\tau}(u) &= u(\tau - I(u < 0))
   \end{aligned}
$$
i.e.  
  
  
$$
\rho_{\tau}(u) = \left\{
        \begin{array}{ll}
            u\tau &, if \quad u \geq 0 \\
            u(\tau - 1) &, if \quad u < 0
        \end{array}
    \right.
$$  
```{r, echo = F}
par(mfrow = c(1, 2))
 plot.new()
 plot.window(xlim = c(-4, 4), ylim = c(-2, 2))
 axis(1, pos = 0, tick = F, labels = F)
 title(main = expression("Graph of loss function for " *  tau *" = 1/2"))
# axis(2, 0:2, pos = c(0, 0), tick = F)
abline(v = 0)
 axis(2, pos = 0, tick = F, labels = F)
 abline(h = 0)
 x1 <- seq(0, 4, length = 1000)
 y <- x1 * .5
 points(x1, y, type = "l", lwd = "2")
 x1 <- seq(-4, 0, length = 1000)
 y <- x1 * -.5
mtext(expression(rho[(1/2)](u)), c(-0.5, 3))
 points(x1, y , type = "l", lwd = "2")
 plot.new()
 plot.window(xlim = c(-4, 4), ylim = c(-2, 2))
 axis(1, pos = 0, tick = F, labels = F)
 # axis(2, 0:2, pos = c(0, 0), tick = F)
 abline(v = 0)
 axis(2, pos = 0, tick = F, labels = F)
 title(main = expression("Graph of loss function for " *  tau *" = 1/3"))
 abline(h = 0)
 x1 <- seq(0, 4, length = 1000)
 y <- x1 * (1/3)
 points(x1, y, type = "l", lwd = "2")
 x1 <- seq(-4, 0, length = 1000)
 y <- x1 * (-2/3)
 points(x1, y , type = "l", lwd = "2")
mtext(expression(rho[(1/3)](u)), c(-0.5, 3))
```
In this graph u is on x-axis and $\rho_{\tau}(u)$ is on y-axis.
For some $\tau \in$ (0, 1) objective is to find $\hat{x}$ which minimizes E[$\rho_{\tau}(X - \hat{x})$].  
$$
\rho_{\tau}(X - \hat{x}) = \left\{
        \begin{array}{ll}
            (X - \hat{x}) \tau &, if \quad (X - \hat{x}) \geq 0 \\
            (X - \hat{x})(\tau - 1) &, if \quad (X - \hat{x}) < 0
        \end{array}
    \right.
$$  
$$
\begin{aligned}
\therefore \quad E[\rho_{\tau}(X - \hat{x})] &= \int_{-\infty}^{\hat{x}}(X - \hat{x})(\tau - 1) dF(x) + \int_{\hat{x}}^{+\infty} (X - \hat{x}) \tau dF(x)\\
&= \tau \int_{-\infty}^{\hat{x}} (X - \hat{x}) dF(x) + \tau \int_{\hat{x}}^{+\infty} (X - \hat{x}) dF(x) - \int_{-\infty}^{\hat{x}} (X - \hat{x}) dF(x)\\
&= \tau \int_{-\infty}^{+\infty} (X - \hat{x}) dF(x) - \int_{-\infty}^{\hat{x}} (X - \hat{x}) dF(x)\\
&= \tau E[X] - \tau \hat{x} - \int_{-\infty}^{\hat{x}} (X - \hat{x}) dF(x) 
\end{aligned}
$$ 

Differentiating both sides w.r.t. $\hat{x}$, we have
$$
 \begin{aligned}
 0 &= 0 - \tau -\int_{-\infty}^{\hat{x}} dF(x)\\
\tau - F(\hat{x}) &= 0\\ 
  F(\hat{x}) = \tau
 \end{aligned}
$$
  
###1.2.2 The Empirical Distribution Function
Suppose X is a real-valued random variable and $(X_1, X_2,......, X_n )$ is a random sample of size n from the distribution of X. Let $(x_1, x_2, ......., x_n)$ be observed values of the random sample.  
The empirical distribution function based on the data $(x_1, x_2, ......., x_n)$ is defined by   
$$
 \begin{aligned}
F_n(x) &= \frac{1}{n} \# \{i \in \{1, 2, ......, n\} : x_i \le x\}\\  
       &= \frac{1}{n} \sum_{i = 1}^{n}[I(x_i \le x)]
 \end{aligned}
$$
In other words if $x_(1) \le x_(2) \le ...... \le x_(n)$, then
$$
 \begin{aligned}
  F_n(x) = \left\{
        \begin{array}{ll}
            0 &,  \quad x \le x_(1) \\
            \frac{1}{n} &, \quad  x_(1) \le x < x_(2) \\
            \frac{2}{n} &, \quad x_(2) \le x < x_(3) \\
            : \\
            :\\
            \frac{n-1}{n} &, \quad x_(n-1) \le x < x_(n)\\
            1 &, \quad x \ge x_(n)
        \end{array}
    \right.
 \end{aligned}
$$  

When F is replaced by the empirical distribution function $F_n(x)$, we may still choose $\hat{x}$ to minimize expected loss:  
        $$
            \begin{aligned}
            \int \rho_{\tau}(x - \hat{x}) dF_n(x) &= \frac{1}{n} \sum_{i = 1}^{n} \rho(x_i - \hat{x})
            \end{aligned}
         $$
 and doing so now yields the $\tau^{th}$ sample quantile.   
 
###1.2.4 Properties of Quantile Functions  
**1**. If F is continuous, strictly increasing on [a, b], then Q is continuous, strictly increasing on [F(a), F(b)] and Q(u) = t iff F(t) = u.  
**Proof**. $inf\{t : F(t) > u^*\} = t^*$  
if $F(t^*) = u^* (since \quad by \quad definition, \quad F^{-1}(u^*) = t*)$  
$\text{This implies that}$ Q(u) = t iff F(t) = u.  
Thus Q on [F(a), F(b)] is the inverse of the restriction of F to [a, b].  
The continuity and increasing character of the inverse is well known.  
  
**2**. If F has jump at t = a, then Q(u) = a, for u $\in$ (F(a-0), F(a)).  
**Proof**.  Since F has jump at t = a.  
$\text{This implies that}$ F(a) $\ge$ u, F(a-0) < u for u $\in$(F(a-0), F(a))  
inf{t:F(t) $\ge$ u} = a.  
Q(u) = a.  
  
**3(i)**. $F^{-1}[F(x)] \le$ x for any x $\in  \mathbb{R}.$  
**(ii)**. $F[F^{-1}(\tau)] \ge \quad \tau$ for any $\tau \in$ (0, 1).  
**Proof(i)**. This follows from the definition .  
$\because \quad F^{-1}[F(x)]$ is the smallest y $\in \mathbb{R}$ with F(y) $\ge$ F(x).  
**(ii)**. This also follows from the definition.  
$F^{-1}(\tau)$ is a value y $\in \mathbb{R}$ satisfying F(y) $\ge \quad \tau$.  
**4**. Quantile function is left continuous i.e. $Q(\tau^{-})$ = $Q(\tau)$.  
**Proof**. This follows from the fact that F is continuous from right.

###1.2.5 Graph of Quantile Function
**Discrete Case **  
Let X be the random variable which takes values 1, 3/2, 2, 5/2, 3 with probability 1/10, 1/5, 3/10, 1/10 respectively.  
Then X has CDF  
$$
 \begin{aligned}
  F(x) = \left\{
        \begin{array}{ll}
            0 &,  \quad x < 1 \\
            \frac{1}{10} &, \quad  1 \le x < \frac{3}{2} \\
            \frac{3}{10} &, \quad \frac{3}{2} \le x < 2 \\
            \frac{6}{10} &, \quad 2 \le x < \frac{5}{2} \\
            \frac{9}{10} &, \quad \frac{5}{2} \le x < 3 \\
            1 &, \quad x \ge 3
        \end{array}
    \right.
 \end{aligned}
$$  

and quantile function 
$$
 \begin{aligned}
  Q(\tau) = \left\{
        \begin{array}{ll}
            1 &,  \quad 0 < \tau \le \quad \frac{1}{10} \\
            \frac{3}{2} &,  \quad \frac{1}{10} < \tau \le \quad \frac{3}{10} \\
            2 &,  \quad \frac{3}{10} < \tau \le \quad \frac{6}{10} \\
            \frac{5}{2} &,  \quad \frac{6}{10} < \tau \le \quad \frac{9}{10} \\
            3 &,  \quad \frac{9}{10} < \tau \le \quad 1 \\
        \end{array}
    \right.
 \end{aligned}
$$  

```{r,echo = FALSE}
par(mfrow = c(1, 2))
 x <- c(1, 3/2, 2, 5/2, 3)
 Fx <- c(1/10, 1/5, 3/10, 3/10, 1/10)
x1 <- c(0, 1/10, 3/10, 6/10, 9/10)
x2 <- c(1/10, 3/10, 6/10, 9/10, 1)
# graph of CDF for the above data
 plot.new()
 plot.window(xlim = c(0, 3.1), ylim = c(0, 1.1))
 axis(1, at = c(1, 3/2, 2, 5/2, 3))
 axis(2, at = c(1/10, 3/10, 6/10, 9/10, 1))
 title(main = "Graph of CDF")
 points(x, x2, pch = 16)
 for(i in 1:(length(x2) - 1)){
     segments(x[i], x2[i], x[i+1], x2[i])
 }
segments(3, 1, 3.3, 1)
 box()
for(i in 1:length(x)){
    abline(v = x[i], lty = "dotted")
}
 
 # grpah of quantile function  for discrete distribution function
 
plot.new()
plot.window(xlim = c(0, 1.2), ylim = c(0, 3.5))
 axis(1, at = c(1/10, 3/10, 6/10, 9/10, 1))
 axis(2, at = c(1, 3/2, 2, 5/2, 3))
 points(x2, x, pch = 16)
 title(main = "Graph of Quantile Function")
 for(i in 1:(length(x1) - 1)){
     segments(x1[i], x[i], x1[i+1], x[i])
 }
segments(9/10, 3, 1, 3)
box()
for(i in 1:length(x2)){
    abline(v = x2[i], lty = "dotted")
}

 

```
  
**Continuous Case**  
Suppose X is a random variable having density function as   
f(x) = $\frac{1}{(x+1)^2}$, x > 0.  
then X has CDF  
$$
\begin{aligned}
F(x) &= \int_{0}^{x} \frac{1}{(t+1)^2} dt\\
&= \frac{-1}{x+1} + 1\\
&= \frac{x}{x+1}, x \ge 0
\end{aligned}
$$ 

Let F(x) = $\tau$, 0 < $\tau$ <1 
$$
\begin{aligned}
\text{This implies that} \quad \frac{x}{x+1} = \tau  \\
 (x+1)\tau = x  \\
 x = \frac{\tau}{1-\tau} \\
\therefore Q(\tau) = \frac{\tau}{1-\tau}, 0 < \tau < 1
\end{aligned}
$$ 
```{r,echo=FALSE}
par(mfrow = c(1, 2))
 curve(x/(x+1), from = 0, to = 100, main = "Graph of CDF", xlab = "x", ylab = "CDF")
curve(x/(1 - x), from = 0, to = 1, main = "Graph of Quantile Funtion", xlab = "tau", ylab = "Q(tau)")

```   
$$
 \begin{aligned}
 \end{aligned}
$$




